Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-01-16 WSSSPE3/report/wg_main_eng_design.tex
Update wg_main_eng_design.tex
Minor additions and deletions

This congregated list is an initial collection of principles that could be extended by adding from further 
related work form separate disciplines within the field of software engineering, including requirements 
engineering, software architecture, and testing. The group identified two example systems to discuss 
the application of the principles. The first one was a quantum chemistry system that allows the analysis 
of the characteristics and capabilities of molecules and solids. The second one was a modeling system 
for malaria that permitted biologists to analyze a range of datasets across geography, biology, and 
epidemiology, and add their own datasets. They then examined the principles and took a retrospective 
analysis of what the developers did in practice against how the principles could have made a 
difference. The opportunity was identified to distill existing software engineering and sustainability 
design knowledge into ``bite sized'' chunks for the Computational Science and Engineering 
Community. In addition, two challenges were pointed out: mapping of the principles to best practice, 
and  demonstrating the return on investment of those best practices.

\subsubsection{Plans}
The next steps in this endeavor are to (1) Systematically analyze a number of example systems from different scientific domains with regards to the identified principles, to (2) Identify the commonalities and gaps in applying those principles to different scientific systems, and to (3) Propose a guideline on the principles and how they exemplary apply to scientific software system. This raised the question, What do the principles mean for scientific and computational software? Similalry, how do the principles relate to non-functional requirements? In addition, what do the five dimensions of sustainability mean for scientific software i.e. environmental, economic, social, techincal and individual?

\subsubsection{Landing Page}
In lieu of a landing page, the Principles for Software Engineering Design for Sustainable Software working group requests an email be sent to Birgit Penzenstadler\footnote{email: \href{mailto:birgit.penzenstadler@csulb.edu}{birgit.penzenstadler@csulb.edu}} and Colin C. Venters\ to find out more about the group's efforts and how to participate.

>>>>

This congregated list is an initial collection of principles that could be extended by adding from further 
related work form separate disciplines within the field of software engineering, including requirements 
engineering, software architecture, and testing. The group identified two example systems to discuss 
the application of the principles. The first one was a quantum chemistry system that allows the analysis 
of the characteristics and capabilities of molecules and solids. The second one was a modeling system 
for malaria that permitted biologists to analyze a range of datasets across geography, biology, and 
epidemiology, and add their own datasets. They then examined the principles and took a retrospective 
analysis of what the developers did in practice against how the principles could have made a 
difference.

\subsubsection{Plans}
The next steps in this endeavor are to (1) Systematically analyze a number of example systems from different scientific domains with regards to the identified principles, to (2) Identify the commonalities and gaps in applying those principles to different scientific systems, and to (3) Propose a set of guidelines on the principles and how they exemplary apply to scientific software system.

\subsubsection{Landing Page}
In the absence of a landing page, the Principles for Software Engineering Design for Sustainable Software working group requests an email be sent to Birgit Penzenstadler\footnote{email: \href{mailto:birgit.penzenstadler@csulb.edu}{birgit.penzenstadler@csulb.edu}} and Colin C. Venters\ to find out more about the group's efforts and how to participate.
